Learning and comparing functional connectomes between populations

### 2012-12-26 paper.tex
Iter on comparison part

\paragraph{Network structure extracted}
%
The correlation matrices and inverse-covariance matrices that we extract
contain a lot of information on the functional structure of the brain.
First, the correlation matrix (Fig.\,\ref{fig:correlation_matrices})
shows blocks of coactivating regions that can be interpreted as
large-scale functional networks, such as the default mode network. Note
the split in network is not straightforward. Different ordering of the
nodes will reveal different networks. Indeed, because of the presence of
hubs and interleaved networks the picture in terms of segregated networks
is not sufficient to explain full-brain connectivity
\cite{varoquaux2012}. The connectivity matrices, correlation matrix or
inverse-covariance matrix, can be represented as graphs: nodes connected
by weighted edges (inserts Fig.\,\ref{fig:correlation_matrices} and
Fig.\,\ref{fig:icov_estimators}). The inverse-covariance matrix,
capturing partial correlations, appears then as extracting a
\emph{backbone} or \emph{core} of the graph. While such structure have
been used as a way to summarize anatomical brain connectivity graphs
\cite{hagmann2008}, here they have a clear-cut meaning with regards to
the BOLD signal: they give the conditional independence structure between
the regions \cite{varoquaux2012}. In other words, regions $a$ and $b$ are
not connected if the signal that they have in common can be explained by
a third region $c$. In this light, the choice of nuisance regressors to
remove confounding common signal is less critical. Indeed, while with
correlation matrices regressing out the global mean has a drastic
(Fig.\,\ref{fig:correlation_matrices} upper right and lower right), on
inverse covariance it only changes very slightly the resulting matrices
(Fig.\,\ref{fig:icov_regressors}). There have been debates on whether to
regress out certain signals, such as the global mean,  as it induces
negative correlations \cite{murphy2009,chang2009}, and these may seem
surprising: one network is doing the opposite of the other. However,
correlation between two signals only takes its meaning with the
definition of a baseline. A simple picture to explain anti-correlations
between two regions is the presence of a third region, mediating the
interactions. Using this third region as a baseline would amount to
estimate partial correlations in the whole system. Using 
inverse-covariance matrices or partial correlations to understand brain
connectivity makes the interpretation in terms of interactions between
brain regions easier and more robust to the choice of confounds.

@@

\section{Comparing connections}

Note here that using a Ledoit-Wolf is probably a good choice for this
application

%------------------------------------------------------------------------------
\subsection{Mass-univariate approaches}

Boils down to a standard linear model => t-tests and co.
Stress that to have Gaussian-distributed correlations, a Fisher Z
transform is required.

Multiple comparison problem is quite an issue (scales a $p^2$) =>
permutation and max-T approach XXX: cite Ge
"resampling-based multiple testing for microarray data analysis", and
Nichols multiple comparison paper.

Family (Cameron, is that the correct name?) FWE?

%------------------------------------------------------------------------------
\subsection{Modeling between-connection dependences}

@@

\section{Comparing Network Summary Statistics}

>>>>

\paragraph{Network structure extracted}
%
The correlation matrices and inverse-covariance matrices that we extract
contain a lot of information on the functional structure of the brain.
First, the correlation matrix (Fig.\,\ref{fig:correlation_matrices})
shows blocks of coactivating regions that can be interpreted as
large-scale functional networks, such as the default mode network. Note
the split in network is not straightforward. Different ordering of the
nodes will reveal different networks. Indeed, because of the presence of
hubs and interleaved networks the picture in terms of segregated networks
is not sufficient to explain full-brain connectivity
\cite{varoquaux2012}. The connectivity matrices, correlation matrix or
inverse-covariance matrix, can be represented as graphs: nodes connected
by weighted edges (inserts on Fig.\,\ref{fig:correlation_matrices} and
Fig.\,\ref{fig:icov_estimators}). The inverse-covariance matrix,
capturing partial correlations, appears then as extracting a
\emph{backbone} or \emph{core} of the graph. While such structure have
been used as a way to summarize anatomical brain connectivity graphs
\cite{hagmann2008}, here they have a clear-cut meaning with regards to
the BOLD signal: they give the conditional independence structure between
the regions \cite{varoquaux2012}. In other words, regions $a$ and $b$ are
not connected if the signal that they have in common can be explained by
a third region $c$. In this light, the choice of nuisance regressors to
remove confounding common signal is less critical. Indeed, while with
correlation matrices regressing out the global mean has a drastic
(Fig.\,\ref{fig:correlation_matrices} upper right and lower right), on
inverse covariance it only changes very slightly the resulting matrices
(Fig.\,\ref{fig:icov_regressors}). There have been debates on whether to
regress out certain signals, such as the global mean,  as it induces
negative correlations \cite{murphy2009,chang2009,fox2009}, and these may seem
surprising: one network is doing the opposite of the other. However,
correlation between two signals only takes its meaning with the
definition of a baseline. A simple picture to explain anti-correlations
between two regions is the presence of a third region, mediating the
interactions. Using this third region as a baseline would amount to
estimate partial correlations in the whole system. Using 
inverse-covariance matrices or partial correlations to understand brain
connectivity makes the interpretation in terms of interactions between
brain regions easier and more robust to the choice of confounds.

@@

\section{Comparing connectivity}

Here, we consider the problem of comparing functional connectivity across
subjects or across conditions.

%------------------------------------------------------------------------------
\subsection{Detecting changing connections}

First, we focus on detecting where the connectivity matrices estimated in
the previous section differ.
\paragraph{Mass-univariate approaches}
%
The most natural approach is to do a linear model on each coefficient of
the connectivity matrices, as in \cite{lewis2009,grillon2012}.
This approach is similar to the second-level
analysis used in mass-univariate brain mapping, and will give rise to
many of the well-known techniques used in such a context, such as the
definition of second-level design, with possibly the inclusion of
confounding effects, and statistical tests (T tests or F tests) on
contrast vector. Importantly, in order to work with Gaussian-distributed
correlations, it is necessary to apply a Fisher Z
transform\footnote{See 
\url{http://en.wikipedia.org/wiki/Fisher_transformation}
or \cite{anderson1958}, section 4.2.3 for mathematical arguments}.
Note that in these settings, the Ledoit-Wolf estimator \cite{ledoit2004}
is probably a good choice to estimate the correlation matrix, as it is
parameter-less and gives good estimation performance without imposing any
restrictions on the data.
%
Correcting for multiple comparison can often severely limit statistical
power, as the number of tests performed scales as the square of the
number of regions used. Controlling for the false discovery rate (FDR)
mitigates this problem. Alternatively, as the assumption underlying the
Benjamini-Holchberg procedure for the FDR can easily be broken,
non-parametric permutations-based approaches give reliable approaches. In
particular, the max-T procedure
% NEEDREF: FDR for neuroimaging data, check the spelling of BH
is interesting to avoid the drastic Bonferroni correction when
controlling for multiple comparison in family-wise error rate.

% XXX: cite Ge "resampling-based multiple testing for microarray data
% analysis", and Nichols multiple comparison paper.

%{{{------ Inter subject fluctuations ----------------------------------------
\begin{figure}
\hspace*{.5ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub00.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub29.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub01.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub07.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub53.pdf}%
\hspace*{.1ex}%
\raisebox{.005\linewidth}{%
    \includegraphics[width=.0186\linewidth]{cmap_00.pdf}%
    \llap{
	\raisebox{.003\linewidth}{%
	    \bfseries\tiny\color{white} -\!1\hspace*{-.2ex}}%
    }%
    \llap{
	\raisebox{.17\linewidth}{%
	    \bfseries\tiny\color{white} 1\hspace*{.1ex}}%
    }%
}%
\llap{\rlap{
    \setlength\fboxsep{1pt}%
    \raisebox{.16\linewidth}{%
	\colorbox{black!10}{%
	    \sffamily\small Covariance matrices}}%
	}%
    \hspace*{.993\linewidth}%
}

\hspace*{.5ex}%
\includegraphics[width=.196\linewidth]{inter_subj/z_score_sub00.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/z_score_sub29.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/z_score_sub01.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/z_score_sub07.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/z_score_sub53.pdf}%
\hspace*{.1ex}%
\raisebox{.005\linewidth}{%
    \includegraphics[width=.0186\linewidth]{cmap_01.pdf}%
    \llap{
	\raisebox{.003\linewidth}{%
	    \bfseries\tiny -\!4\hspace*{-.2ex}}%
    }%
    \llap{
	\raisebox{.168\linewidth}{%
	    \bfseries\tiny 4\hspace*{.1ex}}%
    }%
}%
\llap{\rlap{
    \setlength\fboxsep{1pt}%
    \raisebox{.16\linewidth}{%
	\colorbox{black!10}{%
	    \sffamily\small Z score on difference}}%
	}%
    \hspace*{.993\linewidth}%
}

\hspace*{.5ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub00.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub29.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub01.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub07.pdf}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub53.pdf}%
\hspace*{.1ex}%
\raisebox{.005\linewidth}{%
    \includegraphics[width=.0186\linewidth]{cmap_01.pdf}%
    \llap{
	\raisebox{.003\linewidth}{%
	    \bfseries\tiny -\!4\hspace*{-.2ex}}%
    }%
    \llap{
	\raisebox{.168\linewidth}{%
	    \bfseries\tiny 4\hspace*{.1ex}}%
    }%
}%
\llap{\rlap{
    \setlength\fboxsep{1pt}%
    \raisebox{.16\linewidth}{%
	\colorbox{black!10}{%
	    \sffamily\small Z score on residuals \scriptsize
		\raisebox{.2ex}{\cite{varoquaux2010b}}}}%
	}%
    \hspace*{.993\linewidth}%
}

\caption{Inter subject fluctuations
\label{fig:icov_regressors}
}
\end{figure}
%---------------------------------------------------------------------------}}}

\paragraph{Modeling between-connection dependences}

@@

%------------------------------------------------------------------------------

\subsection{Comparing Network Summary Statistics}
